\documentclass{article}

\input{~/projects/latex/dist/recommended.tex}

\usetikzlibrary{automata, positioning, arrows.meta}
\newcommand{\hdelta}{\hat{\delta}}
\newcommand{\qacc}{q_{\text{accept}}}
\newcommand{\qrej}{q_{\text{reject}}}
\newcommand{\ldiag}{L_{\text{diag}}}
\newcommand{\lempty}{L_{\text{empty}}}
\renewcommand{\tc}{\text{Time}}
\newcommand{\spc}{\text{Space}}

\setup{Theoretical Computer Science - Compact}

\begin{document}
\startDocument
\usetcolorboxes

\vspace{2cm}
\begin{Huge}
    \begin{center}
        TITLE PAGE COMING SOON
    \end{center}
\end{Huge}


\vspace{4cm}
\begin{center}
    \begin{Large}
        ``\textit{Sie können also alle C Programme in Kanonischer Ordnung aufzählen. Sollten Sie dies tun. Wahrscheinlich nicht. Was aber zählt ist, sie \textbf{können} es tun}''
    \end{Large}

    \hspace{3cm} - Prof. Dr. Dennis Komm, 2025
\end{center}

\vspace{3cm}
\begin{center}
    HS2025, ETHZ\\[0.2cm]
    \begin{Large}
        Compact Summary of the book \color{MidnightBlue}\fbox{\href{https://link.springer.com/book/10.1007/978-3-658-06433-4}{Theoretische Informatik}}\color{black}
    \end{Large}\\[0.2cm]
    by Prof. Dr. Juraj Hromkovic
\end{center}


\newpage
\printtoc{Orange}
\section{Introduction}
This summary aims to provide a simple, easy to understand and short overview over the topics covered, with approaches for proofs, important theorems and lemmas,
as well as definitions.

It does not aim to serve as a full replacement for the book or my main summary, but as a supplement to both of them.

It also lacks some formalism and is only intended to give some intuition, six pages are really not enough for a formal and complete overview of the topic.

As general recommendations, try to substitute possibly ``weird'' definitions in multiple choice to see a definition from the book.


\input{parts/01_words-alphabets.tex}
\input{parts/02_finite-automata.tex}
\input{parts/03_turing-machines.tex}
\input{parts/04_computability.tex}
\input{parts/05_complexity.tex}



\end{document}
